---
title: "Flower Species Report: `r params$sp`"
format: html
params:
  sp: "rose"
execute:
  echo: false
  warning: false
  message: false
---

## Overview

This report provides an overview of the flower species: **`r params$sp`**.

It includes a summary table and height distribution plot.

```{r}
library(tidyverse)

# Load dataset
df <- readr::read_csv("data/flower_dataset.csv", show_col_types = FALSE)

# Filter data by species
df_filtered <- df %>% dplyr::filter(species == params$sp)

# Summary statistics
summary_table <- df_filtered %>%
  dplyr::summarise(
    Count = n(),
    Average_Height_cm = round(mean(height_cm), 2),
    Min_Height_cm = min(height_cm),
    Max_Height_cm = max(height_cm)
  )
```

---

## Summary Statistics for `r params$sp`

```{r}
knitr::kable(summary_table, caption = paste("Basic Summary Statistics for", params$sp))
```

**Description:**

The summary table above provides statistics for **`r params$sp`**:

- **Number of records:** `r summary_table$Count`

- **Average height:** `r summary_table$Average_Height_cm` cm

- **Minimum height:** `r summary_table$Min_Height_cm` cm

- **Maximum height:** `r summary_table$Max_Height_cm` cm  

These values offer a numerical overview of this species' height characteristics.

---

## Height Distribution of `r params$sp`

```{r}
# Create the histogram
hist_plot <- ggplot2::ggplot(df_filtered, ggplot2::aes(x = height_cm)) +
  ggplot2::geom_histogram(binwidth = 10, fill = "skyblue", color = "black") +
  ggplot2::labs(
    title = paste("Height Distribution of", params$sp),
    x = "Height (cm)",
    y = "Count"
  ) +
  ggplot2::theme_minimal()

# Save plot with species specific name
hist_path <- paste0("histogram_", gsub(" ", "_", params$sp), ".png")
ggplot2::ggsave(hist_path, plot = hist_plot, width = 6, height = 4)


## Display the histogram
knitr::include_graphics(hist_path)
```

**Description:**

The histogram above visualises how the flower heights for **`r params$sp`** are distributed. The x-axis represents height in centimeters, while the y-axis shows the number of flowers in each bin.

---


## Additional Notes

The species **`r params$sp`** displays a characteristic height range and variety in sizes. These observations can support further analysis of growth patterns, suitability for cultivation, or comparison with other flower species.
